---
title: "Practical Machine Learning Course Final Project"
author: "Paravasthuramesh Velayuthasamy"
date: "28OCT2016"
output: html_document
---


## Overview  
In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

People regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. The goal of this project is to predict the manner in which they did the exercise. 

## Data Preprocessing  
```{r, cache = T, warning = F}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```
  
### Load the dataset
After downloading the data from the data source, we can read the two csv files into two data frames.  
```{r, cache = T,warning = F }
dsTrainingBase <- read.csv("./data/pml-training.csv")
dsTestingBase <- read.csv("./data/pml-testing.csv")
dim(dsTrainingBase)
dim(dsTestingBase)
```
The training DataSet inlcude 19,622 rows 160 variable and the testing DataSet include 20 rows nd 160 variables. 

### Clean the data set - Remove columns [Missing values and not needed]
```{r, cache = T, echo = F}
sum(complete.cases(dsTrainingBase))
dsTrainingBase <- dsTrainingBase[, colSums(is.na(dsTrainingBase)) == 0] 
dsTestingBase <- dsTestingBase[, colSums(is.na(dsTestingBase)) == 0] 
```  
Delete the columns not relevant to the analysis.
```{r, cache = T}

classe <- dsTrainingBase$classe

dsTrainingDelete <- grepl("^X|timestamp|window", names(dsTrainingBase))
dsTrainingBase <- dsTrainingBase[, !dsTrainingDelete]
dsTrainingNew <- dsTrainingBase[, sapply(dsTrainingBase, is.numeric)]

dsTrainingNew$classe <- classe

dsTestDelete <- grepl("^X|timestamp|window", names(dsTestingBase))
dsTestingBase <- dsTestingBase[, !dsTestDelete]
dsTestNew <- dsTestingBase[, sapply(dsTestingBase, is.numeric)]

```
After data clean process we have [19622 rows, 53 variables], [20 rows and 53 variables] with in Training and Test data set respectively. The "classe" variable is still in the cleaned training set.

### Data Slicing
Slice dataset into 70% training and 30% testing which will be used for validataion dataset for Cross validation.

```{r, cache = T}
set.seed(53) 
inTrain <- createDataPartition(dsTrainingNew$classe, p=0.70, list=F)
trainData <- dsTrainingNew[inTrain, ]
testData <- dsTrainingNew[-inTrain, ]
```

## Data Modeling
Random forest model has high accuracy in most case and this model selects most important variables. Also will use 5 fold cross validation with this.
```{r, cache = T}
cvRF <- trainControl(method="cv", 5)
modFitRandForest <- train(classe ~ ., data=trainData, method="rf", trControl=cvRF, ntree=250)
modFitRandForest
```

Then, we estimate the performance of the model on the validation data set.  

```{r, cache = T}
predictRf <- predict(modFitRandForest, testData)
confusionMatrix(testData$classe, predictRf)
```
```{r, cache = T}
modAccuracy <- postResample(predictRf, testData$classe)
modAccuracy
modSampleError <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
modSampleError
```
Estimated accuracy of the model is 99.33% and the estimated sample error is 0.66%.

## Predicting for Test DataSet
Finally prediting on the test dataset.
```{r, cache = T}
answers <- predict(modFitRandForest, dsTestNew[, -length(names(dsTestNew))])
answers
```  
Creat 20 files with the answers. 
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)
```



